Sex and Power: sexual dimorphism in trait variability and its eco-evolutionary and statistical implications

### 2020-04-24 scripts/MouseSexDiffApril20_2020.Rmd
To do: address MISSING comments
naming of chunks, removal of messages

```{r}
library(readr)
library(dplyr)
library(metafor)
library(devtools)
library(purrr)
library(tidyverse)
library(tidyr)
library(tibble)
library(kableExtra)
library(robumeta)
library(ggpubr)
library(ggplot2)
library(here)
# additional packages for intergating male / female symbols in plot 4
library(png)
library(grid)
```

@@

```{r}
calculate_population_stats <- function(mydata, min_individuals = 5) {
  mydata %>%
    group_by(population, strain_name, production_center, sex) %>%
    summarise(
      trait = parameter_name[1],
      x_bar = mean(data_point),
      x_sd = sd(data_point),
      n_ind = n()
    ) %>%
    ungroup() %>%
    filter(n_ind > min_individuals) %>%
    # Check both sexes present & filter those missing
    group_by(population) %>%
    mutate(
      n_sex = n_distinct(sex)
    ) %>%
    ungroup() %>%
    filter(n_sex == 2) %>%
    select(-n_sex) %>%
    arrange(production_center, strain_name, population, sex)
}
```

@@

```{r}
create_meta_analysis_effect_sizes <- function(mydata) {
  i <- seq(1, nrow(mydata), by = 2)
  input <- data.frame(
    n1i = mydata$n_ind[i],
    n2i = mydata$n_ind[i + 1],
    x1i = mydata$x_bar[i],
    x2i = mydata$x_bar[i + 1],
    sd1i = mydata$x_sd[i],
    sd2i = mydata$x_sd[i + 1]
  )

@@

```{r}

@@

This step we have already done and provide a cleaned up file which is less computing intensive and which we have saved in a folder called `export`. However, the cvs is provided in case this is preferred to be attempted, following the steps below: [This code is currently excluded from showing in this html file, but can be viewed and enabled / run in the downloadable .rmd file. However, it requires the .gz file containing the raw data in the "export" folder. Unfortuntaltely, that file can not be provided via github, as is too large (274MB). It will be freely accessible, uploaded on dryad. Link to dryad file HERE]

@@

```{r include=FALSE}
# Initial checks
length(unique(data$parameter_name)) # 232 traits
length(unique(data$parameter_group)) # 161 parameter groups
length(unique(data$procedure_name)) # 26 procedure groups
length(unique(data$biological_sample_id)) # 27147 individial mice  
# To show a table with number of males and females per strain per production center:

@@

```{r}

@@

```{r}

@@

```{r}
results_alltraits_grouping2 <- 
  results_alltraits_grouping %>% 
  left_join(by="id",
             data %>% select(id, parameter_group, procedure = procedure_name, procedure_name, parameter_name) %>%   # We filter duplicated id's to get only one unique row per id (and there is one id per parameter_name)
              filter(!duplicated(id))
            ) %>%
  # Below we add 'procedure' (from the previously loaded 'procedures.csv') as a variable
  left_join(by="procedure", 
            procedures %>% distinct()
            )

@@

```{r}
# Here we double check numbers of trait parameters in the dataset
meta1 <- meta_clean 
# length(unique(meta1$procedure)) #18
# length(unique(meta1$GroupingTerm)) #9
# length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta-analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name"
# length(unique(meta1$parameter_name)) #218
# and prepare the dataset for second-order meta-analysis
meta1_sub <- meta1 %>%
  # Add summary of number of parameter names in each parameter group
  group_by(parameter_group) %>%
  mutate(par_group_size = length(unique(parameter_name)), 
         sampleSize = as.numeric(sampleSize)) %>% 
  ungroup() %>% 
  # Create subsets with > 1 count (par_group_size > 1)
  filter(par_group_size > 1) # 90 observations

@@

```{r}

@@

```{r}
kable(metacombo, digits = 3) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

@@

```{r}
metacombo_final <- metacombo %>%
  group_by(GroupingTerm) %>%
   nest()

@@

```{r}
Behaviour <- overall1 %>% 
  filter(., GroupingTerm == "Behaviour") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,     lnCVR_I2 = .[[4]][[1]]$I2,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se, lnVR_I2 = .[[5]][[1]]$I2,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se, lnRR_I2 = .[[6]][[1]]$I2
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_I2)

@@

```
### Supplemental Table 4
MISSING table descpition. Refer to in ms.
```{r}
kable(overall2, , digits = 3) %>%
 kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

@@

```{r}

@@

```{r}
overall3 <- gather(overall2, parameter, value, c(lnCVR, lnRR), factor_key = TRUE)

@@

# Metameta_Fig3_alltraits
```

@@

```{r}
Fig4 <- ggarrange(malebias_Fig2_alltraits, Metameta_Fig3_alltraits,  nrow = 2, align = "v", heights = c(1, 1), labels = c("A", "B"))
Fig4
```

@@

To do this, we select only traits that have CIs that do not overlap with zero.
The code below creates Figure 5A.
```{r}

@@

```{r}
#Joining the Parts:

@@

The supplemental figures include lnVR (in addition to lnCVR and lnRR, as presented in Figures 4 and 5)
```{r}
# *Prepare data for all traits

@@

```{r}
# Create dataframe to store results
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n * 30), ncol = 30)))
names(results.allhetero.grouping) <- c(
  "id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR",
  "s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR",
  "s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR",
  "s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper",
  "lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper", "lnRR_se"
)

@@

```{r}
results.allhetero.grouping2 <- results.allhetero.grouping[results.allhetero.grouping$s.nlevels.strain.VR != 0, ]
# nrow(results.allhetero.grouping) #232

@@

### Preparation for Figure S2 B (second-order meta-analysis)
#### Female Figure, significant and highly biased traits (Figure S2 B, left hand side panels)
Preparing data for traits with CI not overlapping 0: 
create column with 1= different from zero, 0= zero included in CI (for Figure S2 B, top left panel)
```{r}
#Restructure data for plotting

@@

# Metameta_Fig3_female.perc (Figure S2 bottom left panel)
```
#### Male Figure, significant and highly biased traits (Figure S2 B, right hand side panels)
Restructuring data with significant male bias (for Figure S2 B, top right panel)
```{r}
overall3.male.sigS <- gather(overall.male.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key = TRUE)

@@

```{r}
sessionInfo()
```

>>>>

```{r packages}
library(readr)
library(dplyr)
library(metafor)
library(devtools)
library(purrr)
library(tidyverse)
library(tidyr)
library(tibble)
library(kableExtra)
library(robumeta)
library(ggpubr)
library(ggplot2)
library(here)
# additional packages for intergating male / female symbols in plot 4
library(png)
library(grid)
```

@@

```{r PopStatsFunction}
calculate_population_stats <- function(mydata, min_individuals = 5) {
  mydata %>%
    group_by(population, strain_name, production_center, sex) %>%
    summarise(
      trait = parameter_name[1],
      x_bar = mean(data_point),
      x_sd = sd(data_point),
      n_ind = n()
    ) %>%
    ungroup() %>%
    filter(n_ind > min_individuals) %>%
    # Check both sexes present & filter those missing
    group_by(population) %>%
    mutate(
      n_sex = n_distinct(sex)
    ) %>%
    ungroup() %>%
    filter(n_sex == 2) %>%
    select(-n_sex) %>%
    arrange(production_center, strain_name, population, sex)
}
```

@@

```{r EffectSizeFunction}
create_meta_analysis_effect_sizes <- function(mydata) {
  i <- seq(1, nrow(mydata), by = 2)
  input <- data.frame(
    n1i = mydata$n_ind[i],
    n2i = mydata$n_ind[i + 1],
    x1i = mydata$x_bar[i],
    x2i = mydata$x_bar[i + 1],
    sd1i = mydata$x_sd[i],
    sd2i = mydata$x_sd[i + 1]
  )

@@

```{r MetaAnalysesFunction}

@@

This step we have already done and provide a cleaned up file which is less computing intensive and which we have saved in a folder called `export`. However, the cvs is provided in case this is preferred to be attempted, following the steps below: [This code is currently excluded from showing in this html file, but can be viewed and enabled / run in the downloadable .rmd file. However, it requires the .gz file containing the raw data in the "export" folder. Unfortuntaltely, that file can not be provided via github, as is too large (274MB). It is freely accessible, uploaded on zenodo.org (https://zenodo.org/deposit/3759701)

@@

```{r Table1_prep, include=FALSE}
# Initial checks
length(unique(data$parameter_name)) # 232 traits
length(unique(data$parameter_group)) # 161 parameter groups
length(unique(data$procedure_name)) # 26 procedure groups
length(unique(data$biological_sample_id)) # 27147 individial mice  
# To show a table with number of males and females per strain per production center:

@@

```{r Table1}

@@

```{r loop, message=FALSE, warning=FALSE, include=FALSE}

@@

```{r cleanMeta1}
results_alltraits_grouping2 <- 
  results_alltraits_grouping %>% 
  left_join(by="id",
             data %>% select(id, parameter_group, procedure = procedure_name, procedure_name, parameter_name) %>%   # We filter duplicated id's to get only one unique row per id (and there is one id per parameter_name)
              filter(!duplicated(id))
            ) %>%
  # Below we add 'procedure' (from the previously loaded 'procedures.csv') as a variable
  left_join(by="procedure", 
            procedures %>% distinct()
            )

@@

```{r Table2_Meta1}
# Here we double check numbers of trait parameters in the dataset
meta1 <- meta_clean 
# length(unique(meta1$procedure)) #18
# length(unique(meta1$GroupingTerm)) #9
# length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta-analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name"
# length(unique(meta1$parameter_name)) #218
# and prepare the dataset for second-order meta-analysis
meta1_sub <- meta1 %>%
  # Add summary of number of parameter names in each parameter group
  group_by(parameter_group) %>%
  mutate(par_group_size = length(unique(parameter_name)), 
         sampleSize = as.numeric(sampleSize)) %>% 
  ungroup() %>% 
  # Create subsets with > 1 count (par_group_size > 1)
  filter(par_group_size > 1) # 90 observations

@@

```{r corrTraits}

@@

```{r Table3}
kable(metacombo, digits = 3) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

@@

```{r MetaGrouping}
metacombo_final <- metacombo %>%
  group_by(GroupingTerm) %>%
   nest()

@@

```{r MetaGroup_extraction}
Behaviour <- overall1 %>% 
  filter(., GroupingTerm == "Behaviour") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,     lnCVR_I2 = .[[4]][[1]]$I2,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se, lnVR_I2 = .[[5]][[1]]$I2,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se, lnRR_I2 = .[[6]][[1]]$I2
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_I2)

@@

```
### Supplemental Table 4
MISSING table descpition. Refer to in ms.
```{r Table4}
kable(overall2, , digits = 3) %>%
 kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

@@

```{r Fig4_prep}

@@

```{r Fig4_B}
overall3 <- gather(overall2, parameter, value, c(lnCVR, lnRR), factor_key = TRUE)

@@

#Metameta_Fig3_alltraits
```

@@

```{r Fig4}
Fig4 <- ggarrange(malebias_Fig2_alltraits, Metameta_Fig3_alltraits,  nrow = 2, align = "v", heights = c(8, 8), labels = c("A", "B"))
Fig4
```

@@

To do this, we select only traits that have CIs that do not overlap with zero.
The code below creates Figure 5A.
```{r Fig5_prep}

@@

```{r Fig5}
#Joining the Parts:

@@

The supplemental figures include lnVR (in addition to lnCVR and lnRR, as presented in Figures 4 and 5)
```{r FigS1_prep}
# *Prepare data for all traits

@@

```{r hetero_loop, message=FALSE, warning=FALSE, include=FALSE }
# Create dataframe to store results
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n * 30), ncol = 30)))
names(results.allhetero.grouping) <- c(
  "id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR",
  "s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR",
  "s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR",
  "s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper",
  "lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper", "lnRR_se"
)

@@

```{r hetero}
results.allhetero.grouping2 <- results.allhetero.grouping[results.allhetero.grouping$s.nlevels.strain.VR != 0, ]
# nrow(results.allhetero.grouping) #232

@@

### Preparation for Figure S2 B (second-order meta-analysis)
#### Female Figure, significant and highly biased traits (Figure S2 B, left hand side panels)
Preparing data for traits with CI not overlapping 0: 
create column with 1= different from zero, 0= zero included in CI (for Figure S2 B, top left panel)
```{r FigS2_B}
#Restructure data for plotting

@@

# Metameta_Fig3_female.perc (Figure S2 bottom left panel)
```
#### Male Figure, significant and highly biased traits (Figure S2 B, right hand side panels)
Restructuring data with significant male bias (for Figure S2 B, top right panel)
```{r FigS2_B_right}
overall3.male.sigS <- gather(overall.male.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key = TRUE)

@@

```{r Session}
sessionInfo()
```
